---
title: "Exercise 11 - Spatial Regression"
author: "Robert Windham"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    collapsed: no
    smooth_scroll: yes
    df_print: paged
  pdf_document: default
  html_notebook:
    df_print: paged
    rows.print: 10
    theme: cosmo
    highlight: breezedark
    number_sections: yes
    toc: yes
    toc_float: 
      collapsed: no
      smooth_scroll: yes
editor_options:
  chunk_output_type: inline
  mode: gfm
---



```{r Packages, message=FALSE, warning=FALSE, include=FALSE}
#install.packages('pacman')

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

packages_needed <- c("biscale", "car", "cleangeo", "cowplot", "geosphere", "maps", "mapproj", "maptools", "spatialreg", "spdep", "sf",  "tidyverse","visreg")


sapply(packages_needed, require, character.only=T)

#pk_to_install <- packages_needed [!( packages_needed %in% rownames(installed.packages())  )]
#if(length(pk_to_install)>0 ){
#  install.packages(pk_to_install,repos="http://cra n.r-project.org")
#}



```


```{r import data, include=FALSE}


data <- read.csv('https://raw.githubusercontent.com/chrismgentry/Spatial-Regression/master/Data/childpov18_southfull.csv', 
                 colClasses = c("character", "character", "character", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric"))




```


```{r data fixing, include=FALSE}

names(data)[names(data)=="X2016.child.poverty"] <- "child.pov.2016"




```



```{r data fixing2, include=FALSE}

sc_pov <- data %>% subset(State == "SC")


summary(sc_pov)


```





# Equation


```{r data fixing3, echo=TRUE}


equation <- child.pov.2016 ~ rural + urban + lnlesshs + lnunemployment + lnsinglemom + lnuninsured +
  lnconstruction + lnhealthss + lnretail + lnblack + lnhispanic + lnincome_ratio + lnteenbirth


options(scipen = 5)



ols <- lm(equation, data=sc_pov)
summary(ols)


```







```{r data fixing4, echo=FALSE}

par(mfrow=c(4,4), mar=c(2, 2, 2, 2))
visreg(ols)



```






```{r data fixing5, include=FALSE}

fips <- county.fips
fips.codes <- separate(data = fips, col = polyname, into = c("state", "county"), sep = ",")
sc_fips <- subset(fips.codes, state=="texas", select=fips)

se_shp <- st_read(file.path("se_counties/se_counties.shp"))

sc_sf <- se_shp %>% subset(STATE_NAME == "South Carolina")




```





## Moran's Correlation




```{r data fixing6}


neighb.data <- poly2nb(sc_sf, queen=T)
cont.neighb <- nb2listw(neighb.data,style="W", zero.policy = TRUE)


lm.morantest(ols, cont.neighb)

```



**We can see there is spatial correlation between residuals**





## LaGrange Multiplier Test


```{r data fixing7, echo=FALSE}



lm.LMtests(ols, cont.neighb, test="all")





```


**None are significant, so we would likely want to keep our OLS results**






# Spatially Lagged X Model 



```{r data fixing8, echo=FALSE}


SLX.model <- spatialreg::lmSLX(equation, data=sc_pov, cont.neighb)
summary(SLX.model)


```



## Combined x & lagged x impacts



```{r data fixing9, echo=FALSE}


summary(spatialreg::impacts(SLX.model, cont.neighb), zstats = TRUE)[["pzmat"]]

```





# Spatial Lag Model



```{r data fixing10, echo=FALSE}


sp.lag.model <- spatialreg::lagsarlm(equation, data=sc_pov, cont.neighb)
summary(sp.lag.model, Nagelkerke = TRUE)



```




## Spatial lag mpact Table





```{r data fixing11, echo=FALSE}


summary(spatialreg::impacts(sp.lag.model, listw = cont.neighb, R=100), zstats = TRUE)[["pzmat"]]


```








# Spatial Error Model








```{r data fixing12, echo=FALSE}



sp.err.model <- spatialreg::errorsarlm(equation, data=sc_pov, cont.neighb)
summary(sp.err.model, Nagelkerke = TRUE)


```


Even though our Spatial Error Model is nonsignificant, our significant log variables include less than highschool education, rate of unemployment, single mother households, percent uninsured, and percent of the population that work in construction/retail.





**SLX Model - Adj R-squared = .7311, p-value = .0001274**


**Lag Model - R-squared = .838, p-value = .14104**


**Err Model - R-squared = .84823, p-value = .060513**



**Even though our Error Model is nonsignificant, we are going to use it because it has the highest explanatory power**






# Hausman test




```{r data fixing13, echo=FALSE}


spatialreg::Hausman.test(sp.err.model)


```



**Since it is not significant, we are good to use our error model.**









```{r data fixing14, include=FALSE}



sd.err <- spatialreg::errorsarlm(equation, sc_pov, cont.neighb, etype = "emixed")
sdm <- spatialreg::lagsarlm(equation, sc_pov, cont.neighb, type = "mixed")





```





# Nested Durbin Error Model


```{r data fixing15, echo=FALSE}


summary(sd.err, Nagelkerke = TRUE)


```




## Impacts analysis



```{r data fixing16}


summary(spatialreg::impacts(sd.err, listw = cont.neighb, R = 100), zstats = TRUE)[["pzmat"]]


```

**Not many variables are significant here**




# Comparing models



```{r data fixing17, echo=FALSE}


LR.Sarlm(sd.err, sp.err.model)


```



**Here we would not reject the null hypothesis, so we can restrict our model to a spatial error model.**






```{r data fixing18, include=FALSE}


sc.xy <- st_centroid(sc_sf)

all.dist.k1 <- knn2nb(knearneigh(sc.xy, k=1, longlat = TRUE))

all.dist.k3 <- knn2nb(knearneigh(sc.xy, k=3, longlat = TRUE))

all.dist.k5 <- knn2nb(knearneigh(sc.xy, k=5, longlat = TRUE))

all.max.k1 <- max(unlist(nbdists(all.dist.k1, sc.xy, longlat=TRUE)))

all.max.k3 <- max(unlist(nbdists(all.dist.k3, sc.xy, longlat=TRUE)))

all.max.k5 <- max(unlist(nbdists(all.dist.k5, sc.xy, longlat=TRUE)))


all.sp.dist.k1 <- dnearneigh(sc.xy, d1=0, d2=1 * all.max.k1, longlat = TRUE)

all.sp.dist.k3 <- dnearneigh(sc.xy, d1=0, d2=1 * all.max.k3, longlat = TRUE)

all.sp.dist.k5 <- dnearneigh(sc.xy, d1=0, d2=1 * all.max.k5, longlat = TRUE)


all.dist.neighb.k1 <- nb2listw(all.sp.dist.k1,style="W", zero.policy = TRUE)

all.dist.neighb.k3 <- nb2listw(all.sp.dist.k3,style="W", zero.policy = TRUE)

all.dist.neighb.k5 <- nb2listw(all.sp.dist.k5,style="W", zero.policy = TRUE)

all.dist.lag.k1 <- spatialreg::lagsarlm(equation, data = sc_pov, listw = all.dist.neighb.k1)

all.dist.lag.k3 <- spatialreg::lagsarlm(equation, data = sc_pov, listw = all.dist.neighb.k3)

all.dist.lag.k5 <- spatialreg::lagsarlm(equation, data = sc_pov, listw = all.dist.neighb.k5)







```






# Distance Lag Model




```{r data fixing19, echo=FALSE}

summary(all.dist.lag.k3, Nagelkerke = TRUE)



```


```{r data fixing20, include=FALSE}

all.dist.err.k1 <- spatialreg::errorsarlm(equation, data = sc_pov, listw = all.dist.neighb.k1)
all.dist.err.k3 <- spatialreg::errorsarlm(equation, data = sc_pov, listw = all.dist.neighb.k3)
all.dist.err.k5 <- spatialreg::errorsarlm(equation, data = sc_pov, listw = all.dist.neighb.k5)



```






# Distance Error Model



```{r data fixing21, echo=FALSE}

summary(all.dist.err.k3, Nagelkerke = TRUE)



```



**Distance Lag Model: R-squared = .83, p-value = .933**

**Distance Error Model: R-squared = .84855, p-value = .057019**









```{r data fixing22, include=FALSE}


dist.err.data <- summary(all.dist.err.k3, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output <- cbind.data.frame(sc_pov$FIPS,
                                    dist.err.data$fitted.values, 
                                    dist.err.data$residual, 
                                    sc_pov$child.pov.2016, 
                                    sc_pov$lnsinglemom, 
                                    sc_pov$lnuninsured,
                                    sc_pov$lnunemployment,
                                    sc_pov$lnretail,
                                    sc_pov$lnconstruction,
                                    sc_pov$lnlesshs, 
                                    sc_pov$lnteenbirth,
                                    stringsAsFactors = FALSE)

#Renaming columns
colnames(dist.err.output) <- c("fips","fitted","resid","childpov",
                        "single_mom","uninsured","unemployment","retail","construction","less_hs","teenbirth")




```









```{r data fixing23, include=FALSE}


sc_poly <- merge(x = sc_sf, y = dist.err.output, 
                 by.x = "FIPS", by.y = "fips", all = TRUE)

bivariate_data <- bi_class(sc_poly, x = childpov, y = unemployment, 
                           dim = 3, style = "quantile")

legend <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Unemployment rate",
                    size = 6)





```






```{r data fixing24, include=FALSE}

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))





```





```{r data fixing25, include=FALSE}


unem_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_sf(data = bivariate_data, aes(fill = bi_class), color = "grey50", show.legend = FALSE) +  
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  coord_sf(crs = 4326, xlim=c(-84,-78), ylim=c(31.7,35.5)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Unemployment rate") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))







```





# Map





```{r data fixing26, echo=FALSE}


final_map <- ggdraw() +

draw_plot(unem_pov_map, 0, 0, 1, 1) +

draw_plot(legend, 0.63, 0.035, 0.25, 0.25)

final_map




```














